SMT-Boosted Security Types for Low-Level MPC

### 2025-11-12 toplas26/metaty.tex
checkpoint

Programmers can also specify bindings for local computations, for
example if $\mx{x}{1},\mx{x}{2}$ are shares obtained by, e.g., local
homomorphic addition, the programmer can specify
$[\mx{x}{1},\mx{x}{2}] : \sharety$. This supports compositionality of
type checking. Thus, in the rules for share type annotations of the
form $\notg{\stt} : \sty$, we have different rules for checking local and interactive
$\sharety$ type ascriptions (\TirName{LocalShares} and \TirName{GenShares},
respectively), and for checking local and interactive
$\pubty$ type ascriptions (\TirName{LocalPublic} and \TirName{MaskedPublic},
respectively). Observe the use of definitions $\xdefs$ to refer to
SMT equivalences, and the use of $\Vdash$ to verify share typing.

@@

\begin{definition}[Abstract Share Type Environments]
  We write $\notg{\Gamma}$ to denote sets of bindings $\notg{\stt} : \sty$, and
  $\notg{R}$ to denote sets of expressions (that evaluate to random variables).
  We define evaluation of these, respectively, as follows:
  \begin{mathpar}
  \inferrule
      {\notg{\stt}_1 \redx \stt_1\ \cdots\ \notg{\stt}_n \redx \stt_n}
      {\setit{\notg{\stt}_1 : \sty_1, \ldots, \notg{\stt}_n : \sty_n} \redx
        \setit{\stt_1 : \sty_1, \ldots, \stt_n : \sty_n}}

@@

\begin{definition}[$\metaprot$ Function Types]
  Dependent function types are of the syntactic form
  $\Pi y_1,\ldots,y_n. \notg{\Gamma} \rightarrow (\notg{\Gamma}, \notg{R})$.
  where $y_1,\ldots,y_n$ range over $\stringty$ or $\cidty$ types.
\end{definition}

@@

>>>>

Programmers can also specify bindings for local computations, for
example if $\mx{x}{1},\mx{x}{2}$ are shares obtained by, e.g., local
homomorphic addition, the programmer can specify
$[\mx{x}{1},\mx{x}{2}] : \sharety$. This supports compositionality of
type checking. Thus, in the rules for share type annotations of the
form $\notg{\stt} : \sty$, we have different rules for checking local and interactive
$\sharety$ type ascriptions (\TirName{LocalShares} and \TirName{GenShares},
respectively), and for checking local and interactive
$\pubty$ type ascriptions (\TirName{LocalPublic} and \TirName{MaskedPublic},
respectively). In these rules we use of definitions $\xdefs$ to refer to
SMT equivalences, and $\Vdash$ to verify share typing.

@@

Rules for other commands are defined in Figure \ref{fig-shtripcmd}.
As anticipated above, the \TirName{Defn} rule automatically adds the
protocol definition of variables to $\xdefs$ as a default equivalence,
which may allow automatic share typing without any use of
\TirName{Coerce}. The most complex rules are for function application
and definition (\TirName{App} and \TirName{Fn}). These rules allow for
generalization and instantiation of share type environments and
thus compositionality in typing. Generalized, aka abstract
share type environments are dependent on $\stringty$ and
$\cidty$ types, and we equate them up to evaluation in
our rules.
\begin{definition}[Abstract Share Type Environments]
  We write $\notg{\Gamma}$ to denote sets of bindings $\notg{\stt} : \sty$, and
  $\notg{R}$ to denote sets of expressions (that evaluate to random variables).
  We define evaluation of these, respectively, as follows:
  \begin{mathpar}
  \inferrule
      {\notg{\stt}_1 \redx \stt_1\ \cdots\ \notg{\stt}_n \redx \stt_n}
      {\setit{\notg{\stt}_1 : \sty_1, \ldots, \notg{\stt}_n : \sty_n} \redx
        \setit{\stt_1 : \sty_1, \ldots, \stt_n : \sty_n}}

@@

We also introduce dependent $\Pi$ types for functions, that allow
dependence on $\stringty$ and $\cidty$ types.
\begin{definition}[$\metaprot$ Function Types]
  Dependent function types are of the syntactic form
  $\Pi y_1,\ldots,y_n. \notg{\Gamma} \rightarrow (\notg{\Gamma}, \notg{R})$.
  where $y_1,\ldots,y_n$ range over $\stringty$ or $\cidty$ types.
\end{definition}
As defined in \TirName{Fn}, a dependent function type $\Pi
y_1,\ldots,y_n. \notg{\Gamma_1} \rightarrow (\notg{\Gamma_2},
\notg{R})$ specifies that, given share typing assumptions in
$\notg{\Gamma_1}$, application the function results in the bindings in
$\notg{\Gamma_2}$ by using the random values in $\notg{R}$. It is
always the case that the type is closed, i.e., only $y_1,\ldots,y_n$
can occur free in $\notg{\Gamma_1}$, $\notg{\Gamma_2}$ $\notg{R}$.
Type checking for functions is accomplished by using fresh ``witness''
values of $\stringty$ and $\cidty$ types to instantiate dependent
variables, similarly to the \TirName{GenEntails} rule. As defined in
the \TirName{App} rule, typing ensures that application of the
function instantiates the dependencies, checks that the instantiation
of $\notg{\Gamma_1}$ is contained in the share typing environment, and
adds the instantiation of $\notg{\Gamma_2}$ as an effect, as will as
the instantiation of $\notg{R}$ with linearity enforced.

@@

\subsubsection{Example: GMW \ttt{and} Gate}

Returning to the GMW \ttt{and} gate example introduced in Figure \ref{fig-gmwprepost},
we can extend the definition of \ttt{andgmw} with type annotations to ensure a
valid share typing, as in Figure \ref{fig-gmwsharetypes}. Here a coercion
is needed to specify that the gate output are shares obtained using
$\ttt{r[z]@1}$ as an OTP. The type for $\ttt{andgmw}$ can then be
obtained as:
$$
\ttt{andgmw} : \Pi x,y,z . \varnothing \rightarrow (\setit{[m[z]@1,m[z]@2] : \sharety}, \setit{r[z]@1})  
$$
Note that the function does not require any share type preconditions, since
the (sub)protocol it defines interactively generates new shares using the OTP.
